Interpretable Neural Networks using EAGGA

### 2025-03-15 paper/main.tex
finish MOO

\subsubsection{A-posteriori}
\label{sec-moo-post}
- problem a priori
  * either restrict search space to ``enforce our will'' (e.g. only use 50\% of features), optimise only prediction performance + use this
  * or leave search space unrestricted but adjust loss function to incorporate multiple objectives + take optimum from there
  * in either case: no knowledge of interplay between HP config + performance on all objectives
  * in practical applications, it is oftentimes useful to make the decision of which point of the Pareto set to use a-posteriori, e.g. if only a slight decrease in
    one objective translates to a significant improvement in another that would have been missed if the problem was optimised using a-priori methods
  * a-posteriori evaluates configs just as a-priori, but keeps track of multiple ``best'' (non-dominated) solutions -> makes relationship between HP config + objectives visible
- aside from usual baselines grid and random seach there are multi-objective BO adaptations, mainly using either of two approaches
  (1) fit single surrogate model on scalarised objectives, e.g. ParEGO, which employs the augemented Tchebycheff function as scalarisation to ensure the Pareto front is
      explored sufficiently \citep[pp. 15f]{10.1145/3610536} + \citep[pp. 54-56]{ParEGO}
  (2) fit one surrogate per objective, then
    - either one acquisition function per objective -> return set of promising next candidates
    - or one overall acquisition function aggregating surrogates, e.g. \textit{EHI}, maximises expected improvement of hypervolume \citep[p. 16]{10.1145/3610536} + \citep[pp. 8f]{EHI}
- lastly, there is also multi-objective EA (MOEA) algorithms to explore Pareto front, one of which is NSGA-II (Nondominated Sorting Genetic Algorithm)
  * improves upon predecessor NSGA \citep{NSGA} by making it parameterless, ensuring elitism, and reducing computational complexity of ranking of individuals \citep[p. 182]{NSGA_II}
  * uses all the regular operators, i.e. reproduction, mutation, crossover can be used from single-objective EA (SOEA)
  * difference to SOEA: ranking of individuals (SOEA uses scalar fitness, MOEA mutliple objectives)
    - multi-objective ranking mechanism based on 2 parts
      * non-dominated sorting, ensures elitism
        - iterative procedure to rank individuals in population by their fronts
        (1) determine pareto front, assign rank 0
        (2) remove pareto front from population
        (3) if any individuals left, repeat from 1, increment rank \#
        - \citep[p. 201]{genetic_algos} + \citep[pp. 183f]{NSGA_II}
      * crowding distance, ensures sufficient diversity, i.e. exploration of pareto front
        - assigns score depending on how crowded area around individual in objective space is
        - crowding distance of an individual = mean side length of cuboid spanned by its nearest neighbours as vertices in objective space
        - individuals without two neighbours in a dimension are assigned infinitely high distance value
        - \citep[p. 185]{NSGA_II}
        -> the less the crowding distance, the more ``crowded'' an area is by other individuals
      -> rank individuals by nds front rank (ascending) + crowding distance (descending) as tie breaker
  * ranking used for selecting $\mu$ best individuals to keep for next generation + for reproduction: via binary tournament selection: select two random individuals,
    pick best w.r.t. nds + cd for offspring creation (i.e. mutation / crossover)
  * using nds for ranking / fitness makes intuitive sense, but why cd?
    - goal of MOO: not simply optimise HV, but approx true Pareto front well
    - only having individuals representing one area in objective space makes Pareto front estimate very ``unstable'': removing this area (i.e. solutions form that area)
      would ``collapse'' entire front
    - instead having multiple areas be represented makes the estimate ``stable'': removing one area wouldn't impact Pareto front estimate a lot
    - \citep[p. 185, pp. 189-192]{genetic_algos}

@@

@@

>>>>

\subsubsection{A-posteriori}
\label{sec-moo-post}
The disadvantage of a-priori methods is the missing knowledge of the interplay between a hyperparameter configuration and the trained model's
performance across objectives:
we can either restrict the search space to enforce some will, such as only using a maximum of 50\% of features, optimize for performance and use the
resulting model, or leave the search space unrestricted, adjust the loss function to incorporate all objectives and take its optimum.
In neither case do we know the impact our a-priori trade-off has on the final performance.
In practical applications, the impact on performance across all objectives is what matters.
This is the main advantage of a-posteriori methods: instead of implicitly defining a trade-off prior to training, we still evaluate multiple HP configurations
but keep the set of non-dominated solutions instead of just one that satisfies the trade-off criterion.
This is beneficial as it makes the relationship visible, a practicioner can now see, for instance, that a slight decrease in one objective might translate to
a significant improvement in another that would have been missed if the problem was optimized with a-priori methods.
\\
Aside from the usual baselines grid and random search there are multi-objetive Bayesian Optimization adaptations, mainly using one of two approaches:
\begin{enumerate}
  \item Fitting a single surrogate model on scalarized objectives. \citet[pp. 54-56]{ParEGO} propose \textit{ParEGO}, which employs the augemented Tchebycheff function as
        scalarization to ensure the Pareto front is explored sufficiently.
  \item Fitting one surrogate per objective, then either using one acquisition function per surrogate to return a set of promising candidate query points, or using
        one overall acquisition function to aggregate the surrogates, such as \textit{EHI} as proposed by \citet[pp. 8f]{EHI}, which maximises the expected improvement
        of hypervolume.
\end{enumerate}

Lastly, there is also a multitude of multi-objective evolutionary algorithms (MOEA) to explore the Pareto front.
A key challenge for MOEAs is determining the fitness of an individual, as there are now multiple objectives that need to be incorporated instead of
one as in single-objective EA (SOEA).
We will outline the popular NSGA-II (Nondominated Sorting Genetic Algorithm II), which improves upon its predecessor NSGA \citep{NSGA} by making it
parameterless, ensuring elitism, and reducing the computational complexity of ranking of individuals \citep[p. 182]{NSGA_II}.
NSGA-II uses all the regular operators reproduction, mutation, and crossover that are already known from SOEA.
The difference is in the fitness function, which for NSGA-II comprises two parts: non-dominated sorting and crowding distance.
\\
\textit{Non-dominated sorting}, as described by \citet[p. 201]{genetic_algos} and \citet[pp. 183f]{NSGA_II} is an iterative procedure to ensure
elitist selection through ranking individuals by their fronts.
First, it assigns all individuals of the Pareto front a rank of 0. It then temporarily removes these individuals from the population and repeats
this procedure with an incremented rank number until no individuals are left.
\\
The second part, \textit{crowding distance} ensures sufficient diversity, i.e. exploration of the Pareto front.
The algorithm assigns each individual a score depending on how ``crowded'' the area in objective space around it is. Crowding distance is computed
as the mean side length of the cuboid spanned by its nearest neighbours as vertices in the objective space, where individuals without two neighbours
along some dimension are assigned an infinitely high distance value. The less the crowding distance, the more an area is by other individuals. \citep[p. 185]{NSGA_II}
\\
NSGA-II then ranks individuals by their non-dominated sorting front ranks in ascending order and by their crowding distance in descending order as tie breaker.
This ranking is then used for selecting the $\mu$ best individuals to keep for the next generation.
Reproduction follows a binary tournament selection, where two individuals are randomly drawn from the population and the better one according to the ranking gets
included in the reproduction pool for mutation and crossover.
\\
Conceptually, the reasoning behind using non-dominated sorting to rank individuals is intuitive: the closer an individuals front is to the Pareto front, i.e.
the smaller its front rank, the better it approaches the true Pareto front.
The usage of crowding distance, on the other hand, is not as straightforward. To understand it, we recall the goal of MOO, which is not merely optimizing
hypervolume, but to approximate the true Pareto front well.
Considering the scenario of only having individuals of one area in the objective space included in the Pareto front estimate makes it very unstable: removing this
area (i.e. the solutions from this area) would collapse the entire front estimate.
Considering now the opposite, having multiple areas making up the Pareto front estimate, makes it stable: removing one area would not impact the estimate a lot,
as it's still held up by other individuals. \citep[p. 185, pp. 189-192]{genetic_algos}

@@

  (experimental setup)

@@

  * only 18 datasets
